---
title: "Code_Final_Project"
output: html_document
---
# Data Import
```{r}
# import necessary libraries
library(tidyverse)
library(shiny)
library(plotly)
library(rvest)
library(shinythemes)
library(ggplot2)
library(ggnewscale) # allows for more complex plots

# import base data set
earnings = read.csv('Median Earnings.csv')

earnings
```

# Data Cleaning

Things to clean:
1. Change Column Names
2. Typecast numbers to int
3. Decide if margin of errors want to be kept (might need to make it another column if so) + condense dataframe

```{r}
df <- earnings %>% 
  mutate(X = lag(X, n = 1, default = NA))

females <- df[, 38:ncol(df)] #gets only the columns for females

#adds some of the column names so that we can use the names function
females[3,10] <- "Science and Engineering Related Fields"
females[3,11] <- 'Business'
females[3,12] <- 'Education'
females[3,1] <- 'Total'

#combines the females with the states column
females <- cbind(df[,1], females)
females[3,1] <- 'States'
names(females) <- females[3,]

#gets only the rows and columns wihtout nas
females <- females[5:nrow(females),c(-3,-14)]

#you can change the by to get all of the errors instead
females <- females[seq(1, nrow(df), by = 3),] 

#parses the numbers for all of the columns
for (each in 2:ncol(females)) {
  females[,each] <- parse_number(females[,each])
}

#does the same thing but for males and all genders
males <- df[,20:37]
males[3,10] <- "Science and Engineering Related Fields"
males[3,11] <- 'Business'
males[3,12] <- 'Education'
males[3,1] <- 'Total'
males <- cbind(df[,1], males)
males[3,1] <- 'States'
names(males) <- males[3,]
males <- males[5:nrow(males),c(-3,-14)]
males <- males[seq(1, nrow(df), by = 3),]

for (each in 2:ncol(males)) {
  males[,each] <- parse_number(males[,each])
}

#this one had a lot of columns that were named the wrong thing, so I had to 
#check with the csv and overwrite them
all_gender <- df[,1:19]
all_gender[2,9] <- "Engineering"
all_gender[2,10] <- "Multidisciplinary Studies"
all_gender[2,11] <- "Science and Engineering Related Fields"
all_gender[2,12] <- 'Business'
all_gender[2,13] <- 'Education'
all_gender[2,1] <- 'States'
all_gender[2,2] <- 'Total'
all_gender[2,18] <- 'Communications'
names(all_gender) <- all_gender[2,]
all_gender <- all_gender[5:nrow(all_gender),c(-3,-14)]
all_gender <- all_gender[seq(1, nrow(df), by = 3),]

#parses all of the median incomes into numeric
for (each in 2:ncol(all_gender)) {
  all_gender[,each] <- parse_number(all_gender[,each])
}

#if you want a full df

#adds a row for the gender identifier
females['Gender'] <- rep('Female', nrow(females))
males['Gender'] <- rep('Male', nrow(females))
all_gender['Gender'] <- rep('Everyone', nrow(females))

#binds the rows
full_df <- rbind(females, males, all_gender)
na.omit(full_df) #gets rid of the two rows of nas


#could also merge on state or cbind
merged_df <- merge(females, males, by = 'States') #and you can merge more
```
```{r}
# removes unneeded data frames and variables to free up computer memory
rm(earnings)
rm(df)
rm(each)

#excludes puerto rice and district of columbia since wage data is not present for it
all_gender = all_gender[!(all_gender$States %in% c('Puerto Rico', 'District of Columbia')),]
females = females[!(females$States %in% c('Puerto Rico', 'District of Columbia')),]
males = males[!(males$States %in% c('Puerto Rico', 'District of Columbia')),]
merged_df = merged_df[!(merged_df$States %in% c('Puerto Rico', 'District of Columbia')),]
full_df = full_df[!(full_df$States %in% c('Puerto Rico', 'District of Columbia')),]

# removes NA rows leftover while keeping in actual NA data
all_gender = all_gender[!is.na(all_gender$States),]
females = females[!is.na(females$States),]
males = males[!is.na(males$States),]
merged_df = merged_df[!is.na(merged_df$States),]
full_df = full_df[!is.na(full_df$States),]

# removes Multidisciplinary studies as a comparison option due to NAs. Okay bc vague category anyways
all_gender = subset(all_gender, select = -`Multidisciplinary Studies`)
females = subset(females, select = -`Multidisciplinary Studies`)
males = subset(males, select = -`Multidisciplinary Studies`)
merged_df = subset(merged_df, select = -c(`Multidisciplinary Studies.x`,`Multidisciplinary Studies.y`))
full_df = subset(full_df, select = -`Multidisciplinary Studies`)

# sorts all dataframes by state alphabetically
all_gender = all_gender[order(all_gender$States),]
females = females[order(females$States),]
males = males[order(males$States),]
full_df = full_df[order(full_df$States),]
merged_df = merged_df[order(merged_df$States),]

# reindexes every dataframe. By setting row names to NULL, it reindexes numerically
rownames(all_gender) = NULL
rownames(females) = NULL
rownames(males) = NULL
rownames(full_df) = NULL
rownames(merged_df) = NULL
```

# Question 1: What is the Gender Gap By Major Across The Country?

diverging bar chart: https://r-charts.com/part-whole/diverging-bar-chart-ggplot2/
Up is increase male, down is increase female. Likely won't be much down but that's the point
show equal y-axis on both sides to exacerbate the difference
group my major. Each bar represents a different major. Leave out Multidisciplinary studies due to NAs
Allow grouped bars (dodge) to show up to 3(?) different states for comparison
color by state?
scale_fill_gradient by amount?

```{r}
# Creating Dataframe that shows income difference by state and by major, as well as averaged overall

#gap = 
```

# Question 3:
```{r}
#Adding Livable Wages

website <- read_html("https://worldpopulationreview.com/state-rankings/livable-wage-by-state")

#reads in the livable wages and states column from the websome
income<-website%>%html_nodes(".align-middle:nth-child(3)")%>% html_text()
livable_income <- parse_number(income) #parses to numeric
States <- website%>%html_nodes(".px-2+ .align-middle")%>% html_text()

#combines states and livable wages into one df
livable_wg <- data.frame(States,livable_income)

#merges the states and the data frame that we created with the majors
merged_livable <- merge(full_df, livable_wg, by = 'States')
```

```{r}
#subset with only the median incomes
subset_data <- merged_livable[3:17]

#subtracts the median incomes by the livable wages
all_majors <- subset_data - livable_income

#binds gender and states back onto the df with the majors income
all_majors_state <- cbind("States" = merged_livable$States,"Gender" = 
                            merged_livable$Gender, all_majors)

#pivots the data so that it can be used to create a visualization
long_data <- pivot_longer(all_majors_state, 
                          cols = c("Computers, Mathematics, and Statistics",              
"Biological, Agricultural, and Environmental Sciences",
"Physical and Related Sciences",                       
"Psychology",                                          
"Social Sciences",                                     
"Engineering",                                         
"Multidisciplinary Studies",                           
"Science and Engineering Related Fields",              
"Business",                                            
"Education",                                           
"Literature and Languages",                            
"Liberal Arts and History",                            
"Visual and Performing Arts",                          
"Communications",                                      
"Other" ), 
                          names_to = "Major", 
                          values_to = "Income_Difference")

```

# Shiny UI
```{r}

ui <- fluidPage(
  #creates a theme
    theme = shinytheme("superhero"),
    #page title
    navbarPage("Median Income Examinations by Major and Gender",
        #tab title
        tabPanel("Median Income Compared to Livable Wage by State",
            fluidRow( 
                sidebarPanel(
                    radioButtons(
                        inputId = "gender", #access the widget by "gender"
                        label = "Filter by Gender: ", #title of widget
                        choices = c("Female", "Male", "Everyone"), #choices
                        selected = "Everyone" #Everyone auto selected
                    ),
                    checkboxGroupInput(
                        "majors", #access widget by "majors"
                        label = "Filter by Major", #title of widget
                        choices = unique(long_data$Major), #choices
                        selected = unique(long_data$Major) #selects all majors auto
                    ),
                    selectInput(
                        inputId = "state", #access widget by "state"
                        label = "Filter by State: ", #title of widget
                        choices = unique(long_data$States), #choices
                        selected = 'Alabama' #Selects Alabama auto
                    )
                ),
                mainPanel(
                  #tells shiny to output the plot bargraph
                  plotOutput("bargraph", width = "100%", height = "500px")
                )
            )
        ),
        tabPanel("Drug Satisfaction & Effectiveness Ratings")
    )
)
```

# Shiny Server
```{r}
server <- function(input, output, session) {
  
    #reactive data based on user inputs
    cleaned_data <- reactive({
        #requires gender, majors, and state to exist
        req(input$gender, input$majors, input$state)
      
        #filters the data according to the inputs
        long_data %>%
            filter(Major %in% input$majors, 
                   States == input$state, 
                   Gender == input$gender)
    })

    #render the plot based on filtered data
    output$bargraph <- renderPlot({
        req(cleaned_data())  #ensure data is available
        
        #creates a graph using cleaned_data() and defines x and y axes
        ggplot(cleaned_data(), aes(x = reorder(Major, Income_Difference), y = Income_Difference)) +
            
            #creates a bar graph that changes the fill color according to the number
            geom_bar(stat = "identity", show.legend = FALSE, color = "white", 
                     fill = ifelse(cleaned_data()$Income_Difference < 0, "darkred", "darkgreen")) +
            #changes x and y axes and the title
            xlab("Majors") +
            ylab("Difference Between Major's Median Salary and Livable Wage") +
            labs(title = "How Much More/Less Each Major is Recieving than Livable Wage") +
            #changes the scale to match the min and max income differences
            scale_y_continuous(breaks = seq(min(cleaned_data()$Income_Difference), max(cleaned_data()$Income_Difference), by = 10000),
                               limits = c(min(cleaned_data()$Income_Difference) - 1000, max(cleaned_data()$Income_Difference) + 1000)) +
            coord_flip() +             #flips it to have sidways bar plot
            theme_minimal() + #adds a theme
            theme(plot.title = element_text(hjust = 0.5)) #puts title in the middle
    })
}
```

# Shiny App
```{r}
shinyApp(ui, server)
```